---
output:
  revealjs::revealjs_presentation:
    css: assets/style.css
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  warning=FALSE, message=FALSE, echo=FALSE
)
options(scipen = 999)

library(tidyverse)
library(magrittr)
library(lubridate)
library(here)
library(janitor)
library(refinr)
library(knitr)
library(tsibble)
library(fable)
library(scales)
library(highcharter)
library(tidytext)
library(ggrepel)
library(glue)
```

```{r data, include=FALSE}
fiscal_year <- function(date) {
  if(is.character(date)) as.Date(date)

  ifelse(month(date) >= 10, year(date) + 1, year(date))
}

fpds <- read_csv("raw/radars_extract.csv")

fpds <-
  fpds %>%
  mutate(
    fiscal_year = fiscal_year(action_date),
    quarter = floor_date(action_date, unit = "quarter"),
    ts_quarter = yearquarter(action_date),
    month = floor_date(action_date, unit = "month")
  )

financials <- vars(
  federal_action_obligation,
  current_total_value_of_award,
  base_and_exercised_options_value,
  potential_total_value_of_award,
  base_and_all_options_value
)
```

# radar: a federal market overview {data-background=assets/sweep.gif}

## agenda

- market summary
- key programs
- additional data sources

# market summary

## market size

```{r}

market_size <- fpds %>%
  distinct(award_id_piid, fiscal_year, .keep_all = TRUE) %>%
  group_by(fiscal_year) %>%
  summarise_at(
    financials,
    ~sum(., na.rm = TRUE)
  )

un_derscore <- function(df, var) {
  var <- enquo(var)
  update_var <- quo_name(var)

  df %>%
    mutate(!!update_var := str_replace_all(!!var, "_", " "))
}

p <- market_size %>% gather(key, value, -fiscal_year) %>%
  mutate(value = round(value /1E9, 2)) %>%
  un_derscore(key)

hchart(p, "column", hcaes(fiscal_year, value, group = key))
```

## agencies

```{r}
agencies <- fpds %>%
  distinct(fiscal_year, awarding_sub_agency_name, award_id_piid, .keep_all = TRUE) %>%
  group_by(fiscal_year, awarding_sub_agency_name) %>%
  summarise_at(
    financials,
    ~sum(., na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(fiscal_year, desc(current_total_value_of_award)) %>%
  filter(current_total_value_of_award > 0) %>%
  rename(`federal action obligation` = federal_action_obligation)

hchart(agencies, "column", hcaes(fiscal_year, `federal action obligation`, group = awarding_sub_agency_name)) %>%
  hc_xAxis(title = "federal action obligation")
```

## major vendors

```{r}

refine <- function(df, col) {
  col <- enquo(col)
  refined_col <- paste0("refined_", quo_name(col))

  new_col <- df %>%
    pull(!!col) %>%
    refinr::key_collision_merge() %>%
    refinr::n_gram_merge()

  df %>%
    mutate(
      !!refined_col := new_col
    )
}

awardees <- fpds %>%
  mutate(
    simple_recip_name = case_when(
      str_detect(recipient_name, "L(-)?3") ~ "L-3",
      str_detect(recipient_name, "GENERAL DYNAMICS") ~ "GENERAL DYNAMICS",
      str_detect(recipient_name, "HONEYWELL") ~ "HONEYWELL",
      str_detect(recipient_name, "LOCKHEED MARTIN") ~ "LOCKHEED MARTIN",
      str_detect(recipient_name, "RAYTHEON") ~ "RAYTHEON",
      str_detect(recipient_name, "ROCKWELL COLLINS") ~ "ROCKWELL COLLINS",
      str_detect(recipient_name, "BOEING") ~ "BOEING",
      str_detect(recipient_name, "BAE SYSTEMS") ~ "BAE SYSTEMS",
      str_detect(recipient_name, "GE AVIATION") ~ "GE AVIATION",
      str_detect(recipient_name, "NORTHROP GRUMMAN") ~ "NORTHROP GRUMMAN",
      str_detect(recipient_name, "LEIDOS") ~ "LEIDOS",
      TRUE ~ recipient_name
    )
  ) %>%
  refine(simple_recip_name) %>%
  group_by(fiscal_year, refined_simple_recip_name) %>%
  summarise_at(financials, ~sum(., na.rm = TRUE)) %>%
  ungroup()

awards <- awardees %>%
  group_by(fiscal_year) %>%
  top_n(10, federal_action_obligation) %>%
  group_by(refined_simple_recip_name, fiscal_year) %>%
  mutate(total_obl = sum(federal_action_obligation)) %>%
  ungroup() %>%
  arrange(desc(total_obl)) %>%
  rename(
    awardee = refined_simple_recip_name,
    `federal action obligation` = federal_action_obligation
  )

hchart(awards, "scatter", hcaes(awardee, `federal action obligation`, group=fiscal_year)) %>%
  hc_chart(inverted = TRUE)
```

## quarterly psc breakdown

```{r}
psc <-
  fpds %>%
  group_by(product_or_service_code_description, quarter) %>%
  summarise_at(financials, ~sum(., na.rm = TRUE)) %>%
  rename(
    `federal action obligation` = federal_action_obligation
  )

hchart(
  psc,
  "column",
  hcaes(quarter, `federal action obligation`, group = product_or_service_code_description)
)

```

## forecast

```{r}

quarterly_obligations <- fpds %>%
  select(ts_quarter, product_or_service_code, federal_action_obligation) %>%
  group_by(ts_quarter, product_or_service_code) %>%
  summarise(obligations = sum(federal_action_obligation, na.rm = TRUE)) %>%
  as_tsibble(key = id(product_or_service_code), index= ts_quarter)


obs <- quarterly_obligations %>%
  # fill_gaps(obligations = 0) %>%
  # fill(obligations) %>%
  model(
    ets = ETS(log(obligations))
  )

obs %>%
  forecast(h=8) %>%
  autoplot(as_tsibble(quarterly_obligations)) +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_yearweek() +
  labs(x = "quarter")
```

# programs

## key programs

```{r}
progs <-
  fpds %>%
  replace_na(list(dod_acquisition_program_description = "NONE")) %>%
  mutate(
    mdap = case_when(
      dod_acquisition_program_description == "NONE" ~ "other",
      TRUE ~ "mdap"
    )
  )

# spend_by_prog <-
mdap_scatter <- progs %>%
  group_by(dod_acquisition_program_code) %>%
  summarise(
    program = first(dod_acquisition_program_description),
    federal_action_obligation = sum(federal_action_obligation, na.rm = TRUE),
  ) %>%
  filter(federal_action_obligation != 0)

mdap_scatter %>%
  ggplot(aes(reorder(program, federal_action_obligation), federal_action_obligation)) +
  geom_point(color = "#f4772a") + coord_flip() +
  labs(x="") +
  scale_y_continuous(labels = scales::dollar) +
  theme_minimal()

```

```{r, results="asis", include=FALSE}
non_mdap <- progs %>%
  filter(dod_acquisition_program_description == "NONE") %>%
  select(awarding_sub_agency_name, award_description, federal_action_obligation, base_and_exercised_options_value)

custom_stopwords <- tibble(word = c("modification", "radar"))

non_mdap_tidy <-
  non_mdap %>%
  unnest_tokens(word, award_description) %>%
  anti_join(stop_words) %>%
  anti_join(custom_stopwords)

tfidf <- non_mdap_tidy %>%
  group_by(awarding_sub_agency_name, word) %>%
  summarize(
    n = n(),
    median_action_value = mean(base_and_exercised_options_value, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  bind_tf_idf(word, awarding_sub_agency_name, n )

top_tf_idf <- tfidf %>%
  group_by(awarding_sub_agency_name) %>%
  top_n(50, tf_idf)

word_value <- function(x) {
  p <- top_tf_idf %>%
    drop_na(word) %>%
    filter(awarding_sub_agency_name == x) %>%
    ggplot(aes(median_action_value, tf_idf, label = word)) +
    geom_text_repel(
      segment.alpha = 0,
      aes(colour=median_action_value)
    ) +
    # scale_colour_viridis_c(option = "C") +
    labs(
      x = "Average contract value (by word)",
      y = "Word importance"
    ) +
    theme_minimal() +
    guides(color = "none") +
    scale_x_continuous(labels = scales::dollar)

  print(p)
}
```

## "chatterplots:"

- compare the importance of a word in its context with the average contract value of what the word describes

## navy

```{r}
word_value("DEPT OF THE NAVY")
```

## army

```{r}
word_value("DEPT OF THE ARMY")
```

## air force

```{r}
word_value("DEPT OF THE AIR FORCE")
```

# additional data sources

## DOD comptroller

- budget requests for increases in programs with heavy radar component
- [data](https://comptroller.defense.gov/Budget-Materials/)

## new solicitations

- program office solicitations for radar and related systems
- [data](ftp://ftp.fbo.gov/)

## selected acquisition reports

- selected acquisition reports for major defense acquisition programs
- [data](https://www.acq.osd.mil/ara/am/sar/)

## commercial innvoations

- commercial developments in sectors with possible spillover benefits
- [IQT Portfolio](https://www.iqt.org/portfolio/)

# questions? {data-background=assets/radar.gif}
